---
title: "Homework Assignment #4"
subtitle: "Identifying Suitable EEZs for Oysters and Pacific Herring"
author: "Ryan Green"
date: 11/30/24
execute:
  warning: false
  message: false
  results: hide
format:
  html:
    toc: true
editor_options: 
  chunk_output_type: console
---

For this assignment, I will identify the most suitable Exclusive Economic Zones (EEZ) on the West Coast of the US for developing marine aquaculture for oysters and Pacific herring (*Clupea pallasii*). Ideal EEZs will be determined based on the range of suitable Sea Surface Temperatures (SST) and depth ranges for each. 


## Initial Data Loading and Manipulation

```{r, Loading Libraries}
# Loading Libraries
library(sf) 
library(tmap)
library(here)
library(terra)
library(stars)
library(dplyr)
library(raster)
library(tidyverse)
library(RColorBrewer)
```

```{r, Loading Data, results='hide'}
# Loading in all data
depth <- rast(here::here("data", "depth.tif"))

eez <- st_read(here::here("data", "wc_regions_clean.shp"))

sst08 <- rast(here::here("data", "average_annual_sst_2008.tif"))
sst09 <- rast(here::here("data", "average_annual_sst_2009.tif"))
sst10 <- rast(here::here("data", "average_annual_sst_2010.tif"))
sst11 <- rast(here::here("data", "average_annual_sst_2011.tif"))
sst12 <- rast(here::here("data", "average_annual_sst_2012.tif"))
```

```{r, CRS Transformation, results='hide'}
# Reprojecting all rasters to the same CRS
sst08 <- terra::project(sst08, "EPSG:4326")
sst09 <- terra::project(sst09, "EPSG:4326")
sst10 <- terra::project(sst10, "EPSG:4326")
sst11 <- terra::project(sst11, "EPSG:4326")
sst12 <- terra::project(sst12, "EPSG:4326")
elevation <- terra::project(depth, "EPSG:4326")
eez <- st_transform(eez, crs = 4326)
```

```{r, CRS Check}
# Checking and converting CRS of SST rasters
rasters <- list(sst08, sst09, sst10, sst11, sst12, elevation)
names(rasters) <- c("sst08", "sst09", "sst10", "sst11", "sst12", "elevation")

for (i in seq_along(rasters)) {
  if (st_crs(rasters[[i]]) != st_crs(4326)) {
    message(paste("Reprojection of", names(rasters)[i], "FAILED"))
  } else {
    message(paste("CRS of", names(rasters)[i], "CONFIRMED as EPSG:4326"))
  }
}

if (st_crs(eez)$epsg == 4326) {
  message("CRS of eez CONFIRMED as EPSG:4326")
} else {
  message("Reprojection of 'eez' FAILED")
}
```

```{r, Calculate and Resample Mean SST}
# Stacking the SST rasters together
sst_stack <- rast(list(sst08, sst09, sst10, sst11, sst12))

# Finding the mean SST from the raster stack, and converting from Kelvin to Celsius
mean_sst <- mean(sst_stack, na.rm = TRUE) - 273.5

# Cropping the elevation/bathymetry raster to the same extent as the SST raster
elevation <- crop(elevation, ext(mean_sst))

# Resampling the elevation raster based on mean_sst
elevation <- resample(elevation, mean_sst, method = "near")
```

## Elevation and Bathymetry Map

```{r, Bathymetry, Elevation Map Setup}
# Creating custom breaks for elevations above and below zero, for color mapping
below_zero_breaks <- c(-6000, -5500, -5000, -4500, -4000, -3500, -3000, -2500, -2000, -1500, -1000, -500, -400, -300, -200, -100, 0)
above_zero_breaks <- c(0, 25, 50, 75, 100, 200, 300, 400, 500, 1000, 1500, 2000, 2500, 3000, 3500, 4000, 4500, 5000, 5500, 5600, 7000)

# Setting the colors
below_zero_palette <- colorRampPalette(c("#07073D", "royalblue2"))(length(below_zero_breaks) - 1)
above_zero_palette <- colorRampPalette(c("#5C653F", "#323615", "aliceblue"))(length(above_zero_breaks) - 1)

# Combining the breaks into a single object
breaks <- c(below_zero_breaks, above_zero_breaks[-1])

# Combining the palette into a single object
palette <- c(below_zero_palette, above_zero_palette)
```

```{r, Map of Bathymetry and Elevation, fig.height=8.7, fig.width=7.9, fig.align='center'}
#| code-fold: TRUE 
#| collapse: TRUE

tm_shape(elevation) +
  tm_raster(palette = palette, 
            breaks = breaks, 
            title = "Elevation (m)", 
            style = "fixed",
            legend.reverse = TRUE) +
tm_layout(legend.outside = TRUE,
          frame = TRUE,
          bg.color = "grey12",
          main.title = "West Coast Elevation and Bathymetry",
          main.title.position = c(0.14, 1),
          main.title.size = 1.2,
          inner.margins = c(0, 0, 0, 0),
          outer.margins = c(0, 0, 0, 0),
          asp = 0) +
tm_scale_bar(text.color = "#F6EAC7",
             color.dark = "#665A47",
             color.light = "#F6EAC7",
             position = c(0.07, 0.02),
             width = 0.3) +
tm_compass(type = "arrow",
           size = 1.5,
           position = c(0.06, 0.06),
           text.color = "#F6EAC7",
           color.dark = "#665A47",
           color.light = "#F6EAC7") 
```

```{r, Map of Mean SST Setup}
# Making a copy of elevation, just for land area
land <- elevation

# Isolating land area by removing any elevation below 0
land[land <= 0] <- NA

# Setting a custom palette for land elevation, this time in grey
above_zero_palette_greys <- colorRampPalette(c("grey40", "grey90"))(length(above_zero_breaks) - 1)

# Creating custom breaks and color palette for the sea surface temperature
temp_breaks <- seq(from = 0, to = 20, by = 1)
temp_palette <- colorRampPalette(c("aliceblue", "darkred"))(length(temp_breaks) - 1)
```

## Sea Surface Temperature Map

```{r, Map of Mean SST, fig.height=8.7, fig.width=7.8, fig.align='center'}
#| code-fold: TRUE 
#| collapse: TRUE

tm_shape(land) +
  tm_raster(palette = above_zero_palette_greys, 
            breaks = above_zero_breaks, 
            title = "Elevation (m)", 
            style = "fixed",
            legend.show = FALSE) +
tm_shape(mean_sst) +
  tm_raster(palette = temp_palette, 
            breaks = temp_breaks, 
            title = "Mean SST (C)",
            legend.reverse = TRUE) +
tm_layout(legend.outside = TRUE,
          frame = TRUE,
          bg.color = "grey40",
          main.title = "West Coast Mean Sea Surface Temperature",
          main.title.position = c(0.14, 1),
          main.title.size = 1.2,
          inner.margins = c(0, 0, 0, 0),
          outer.margins = c(0, 0, 0, 0),
          asp = 0) +
tm_scale_bar(text.color = "aliceblue",
             color.dark = "black",
             color.light = "aliceblue",
             position = c(0.07, 0.02),
             width = 0.3) +
tm_compass(type = "arrow",
           size = 1.5,
           position = c(0.06, 0.06),
           text.color = "aliceblue",
           color.dark = "black",
           color.light = "aliceblue") 
```


## Oyster Aquaculture

```{r, Oyster Temp}
# Making a copy of mean_sst, just for our oysters :)
oyster_temp <- mean_sst

# Establishing reclassification matrix for oyster_temp
rcl_temp <- matrix(c(-Inf, 11, 1,
                11, 30, 0,
                30, Inf, 1), 
                ncol = 3, byrow = TRUE)

# Reclassifying oyster_temp based on the matrix
oyster_temp <- classify(oyster_temp, rcl = rcl_temp)

# Overwriting oyster_temp values as factor
oyster_temp <- as.factor(oyster_temp)
```

```{r, Oyster Depth}
# Just like oyster_temp, copying the elevation raster for oyster_depth
oyster_depth <- elevation

# Establishing reclassification matrix for oyster_depth
rcl_depth_fun <- matrix(c(-Inf, -70, 0,
                -70, 0, 1,
                0, Inf, 0), 
                ncol = 3, byrow = TRUE)

# Reclassifying oyster_depth based on the matrix
oyster_depth <- classify(oyster_depth, rcl = rcl_depth_fun)

# Overwriting oyster_depth values as factor
oyster_depth <- as.factor(oyster_depth)
```

```{r, Oyster Checks}
# Running CRS, Extent, Resolution, and NA values checks for oyster_temp and oyster_depth
checks <- list(
  Extent_Check = ext(oyster_temp) == ext(oyster_depth),
  Resolution_Check = res(oyster_temp) == res(oyster_depth),
  CRS_Check = crs(oyster_temp) == crs(oyster_depth),
  NaN_Temp_Check = !any(is.nan(values(oyster_temp))),
  NaN_Depth_Check = !any(is.nan(values(oyster_depth)))
)

for (check_name in names(checks)) {
  check_result <- checks[[check_name]]
  print(paste(check_name, ":", check_result))
}

if (all(c(checks$ext_check, checks$res_check, checks$crs_check) == TRUE) &
    all(c(checks$nan_check_temp, checks$nan_check_depth) == TRUE)) {
  print("Checks passed")
} else {
  print("Checks failed")
}
```

```{r, Oyster Territory Raster}
# Finding suitable oyster territory based on depth and temp
oyster_territory <- (oyster_depth * oyster_temp)

# Saving as factor
oyster_territory <- as.factor(oyster_territory)

# Setting 0 values to NA, to only get the suitable area in the oyster_territory raster
values(oyster_territory)[values(oyster_territory) == 0] <- NA

# Transforming eez to the same CRS as oyster_territory
eez <- st_transform(eez, crs = crs(oyster_territory))

# Rasterizing eez
eez_raster <- rasterize(eez, oyster_territory, field = "rgn_id")

# Finding the suitable territory within each eez
oyster_territory <- (oyster_territory * eez_raster)
```

## Oyster Aquaculture EEZ Suitability Table

```{r, Oyster Territory Info}
# Reprojecting oyster_territory to a CRS that uses square meters, rather than degrees
oyster_territory <- project(oyster_territory, "EPSG:3395")

# Finding resolution of oyster_territory_m
cell_size <- res(oyster_territory)

# Calculating area of a single cell
cell_area_m2 <- prod(cell_size)

# Converting to square KM
cell_area_km2 <- cell_area_m2 / 1000000

# Getting the count of each unique value in oyster_territory_m
value_counts_oyster <- freq(oyster_territory)

# Calculating total area for each value
value_counts_oyster$area_km2_o <- value_counts_oyster$count * cell_area_km2

# Renaming column referring to EEZ to "rgn_id", so that I can join the region names to this dataframe
colnames(value_counts_oyster)[colnames(value_counts_oyster) == "value"] <- "rgn_id"

# Dropping eez geometry to make it a dataframe
region_info <- eez %>%
  st_drop_geometry()

# Joining region_info with value_counts_oyster to add region names to the dataframe
value_counts_oyster <- region_info %>%
  left_join(value_counts_oyster, by = "rgn_id") %>%
  dplyr::select(-area_m2, -area_km2, -layer, -rgn_key) %>%
  mutate(across(everything(), ~replace(., is.na(.), 0)))

# Joining the oyster value counts to eez_all, a dataframe used for making map labels
eez_all <- eez %>%
  left_join(value_counts_oyster, by = "rgn_id")

# Creating combined map labels with region name and code
eez_all$combined_labels <- paste(eez_all$rgn.y, eez_all$rgn_key, sep = "\n")

# Making a nice kable table, nice!
kableExtra::kable(value_counts_oyster, 
                  format = "simple", 
                  col.names = c("Region", 
                                "Region ID", 
                                "Cell Count", 
                                "Area, Km²"), 
                  caption = "Suitable Oyster Cultivation Territory in West Coast EEZs",
                  align = "l",
                  digits = 2)

```

## Oyster Aquaculture EEZ Map

```{r, Oyster Aquaculture EEZ Map, fig.height=9, fig.width=5.5, fig.align='center'}
#| code-fold: TRUE 
#| collapse: TRUE

tm_shape(elevation) +
  tm_raster(palette = palette, 
            breaks = breaks, 
            title = "Elevation (m)", 
            style = "fixed",
            alpha = 0.8,
            legend.show = FALSE) +
tm_shape(eez_all) +
  tm_polygons(col = "area_km2_o",
              palette = "Greens",
              border.col = "#6A8EAE",
              lwd = 0.3,
              alpha = 0.9,
              title = expression("Suitable Area, Km"^2*"")) +
  tm_text("combined_labels", 
          size = 0.5, 
          col = "midnightblue", 
          shadow = FALSE, 
          auto.placement = FALSE) + 
tm_layout(legend.text.size = 0.6,
          legend.width = 0.3,
          legend.position = c(0.72, 0.924),
          legend.bg.color = "seashell1",
          frame = TRUE,
          bg.color = "grey12",
          main.title = "West Coast EEZ Suitability for\nOyster Aquaculture",
          main.title.position = c("center", "top"),
          main.title.size = 1.2,
          inner.margins = c(0, 0, 0, 0),
          outer.margins = c(0, 0, 0, 0),
          asp = 0) +
tm_scale_bar(text.color = "#F6EAC7",
             color.dark = "#665A47",
             color.light = "#F6EAC7",
             position = c(0.07, 0.02),
             width = 0.3) +
tm_compass(type = "arrow",
           size = 1.5,
           position = c(0.06, 0.06),
           text.color = "#F6EAC7",
           color.dark = "#665A47",
           color.light = "#F6EAC7") 
```


## Herring Aquaculture
Finding suitable territory for Herring, using the same steps as the oysters. This code is nearly identical to that of the oysters, so I have omitted comments to keep it neat. 

```{r, Herring Temp}
herring_temp <- mean_sst
herring_temp[is.nan(herring_temp)] <- 0

# Slightly different matrix, as herring have different preferred temperatures
rcl_temp_h <- matrix(c(-Inf, 0, 0,
                0, 10, 1,
                10, Inf, 0), 
                ncol = 3, byrow = TRUE)

herring_temp <- classify(herring_temp, rcl = rcl_temp_h)

herring_temp <- as.factor(herring_temp)
```

```{r, Herring Depth}
herring_depth <- elevation

# Slightly different matrix, as herring have different preferred depths
rcl_depth_fun_h <- matrix(c(-Inf, -1300, 0,
                -1300, 0, 1,
                0, Inf, 0), 
                ncol = 3, byrow = TRUE)

herring_depth <- classify(herring_depth, rcl = rcl_depth_fun_h)

herring_depth <- as.factor(herring_depth)
```

```{r, Herring Territory Raster}
herring_territory <- (herring_depth * herring_temp)

herring_territory <- as.factor(herring_territory)

values(herring_territory)[values(herring_territory) == 0] <- NA

eez <- st_transform(eez, crs = crs(herring_territory))

eez_raster <- rasterize(eez, herring_territory, field = "rgn_id")

herring_territory <- (herring_territory * eez_raster)
```

## Herring Aquaculture EEZ Suitability Table

```{r, Herring Territory Info}
herring_territory <- project(herring_territory, "EPSG:3395")

cell_size <- res(herring_territory)

cell_area_m2 <- prod(cell_size)

cell_area_km2 <- cell_area_m2 / 1000000

value_counts_herring <- freq(herring_territory)

value_counts_herring$area_km2_h <- value_counts_herring$count * cell_area_km2

colnames(value_counts_herring)[colnames(value_counts_herring) == "value"] <- "rgn_id"

region_info <- eez %>%
  st_drop_geometry()

value_counts_herring <- region_info %>%
  left_join(value_counts_herring, by = "rgn_id") %>%
  dplyr::select(-area_m2, -area_km2, -layer, -rgn_key) %>%
  mutate(across(everything(), ~replace(., is.na(.), 0)))

eez_all <- eez_all %>%
   left_join(value_counts_herring, by = "rgn_id")
 
kableExtra::kable(value_counts_herring, 
                  format = "simple", 
                  col.names = c("Region", 
                                "Region ID", 
                                "Cell Count", 
                                "Area, Km²"), 
                  caption = "Suitable Herring Cultivation Territory in West Coast EEZs",
                  align = "l",
                  digits = 2)
```

## Herring Aquaculture EEZ Map

```{r, Herring Aquaculture EEZ Map, fig.height=9, fig.width=5.5, fig.align='center'}
#| code-fold: TRUE 
#| collapse: TRUE

tm_shape(elevation) +
  tm_raster(palette = palette, 
            breaks = breaks, 
            title = "Elevation (m)", 
            style = "fixed",
            alpha = 0.8,
            legend.show = FALSE) +
tm_shape(eez_all) +
  tm_polygons(col = "area_km2_h",
              palette = "Greens",
              border.col = "#6A8EAE",
              lwd = 0.3,
              alpha = 0.9,
              title = expression("Suitable Area, Km"^2*"")) +
  tm_text("combined_labels", 
          size = 0.5, 
          col = "midnightblue", 
          shadow = FALSE, 
          auto.placement = FALSE) + 
tm_layout(legend.text.size = 0.6,
          legend.width = 0.3,
          legend.position = c(0.72, 0.884),
          legend.bg.color = "seashell1",
          frame = TRUE,
          bg.color = "grey12",
          main.title = "West Coast EEZ Suitability for\nHerring Aquaculture",
          main.title.position = c("center", "top"),
          main.title.size = 1.2,
          inner.margins = c(0, 0, 0, 0),
          outer.margins = c(0, 0, 0, 0),
          asp = 0) +
tm_scale_bar(text.color = "#F6EAC7",
             color.dark = "#665A47",
             color.light = "#F6EAC7",
             position = c(0.07, 0.02),
             width = 0.3) +
tm_compass(type = "arrow",
           size = 1.5,
           position = c(0.06, 0.06),
           text.color = "#F6EAC7",
           color.dark = "#665A47",
           color.light = "#F6EAC7") 
```


# Function

```{r}


```







